---
title: "imdbtop1000project(part2)"
output: html_document
date: "2023-09-27"
---
```{r include=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(janitor)
library(ggplot2)
library(ggthemes)

imdbtop1000 <-read.csv("~/Downloads/imdb_top_1000.csv", sep = ",")

imdbtop1000v1 <- imdbtop1000%>%
  clean_names()

imdbtop1000v1 %>%
  get_dupes(poster_link)

summary(imdbtop1000v1)

imdbtop1000v2 <- imdbtop1000v1 %>%
  mutate(meta_score = coalesce(meta_score, 78))

summary(imdbtop1000v2)

imdbtop1000v2 %>%
  count(released_year)

imdbtop1000v2%>%
  filter(released_year == 'PG')

imdbtop1000v4 <- imdbtop1000v2 %>% 
  mutate(released_year = str_replace(released_year, 'PG', '1995'))

imdbtop1000v4 <- imdbtop1000v4 %>%
  mutate(certificate = str_replace(certificate, 'U', 'PG'))

imdbtop1000v4%>%
  filter(series_title == 'Apollo 13')

imdbtop1000v4%>%
  count(certificate)

imdbtop1000v4%>%
  filter(certificate == 16)

imdbtop1000v4%>%
  filter(certificate == 'PGnrated')

imdbtop1000v4 %>%
  count(certificate)


imdbtop1000v7<- imdbtop1000v4%>%
  mutate(certifcate = str_replace(certificate, "NA", "None"))

imdbtop1000v7 <- imdbtop1000v4%>%
  mutate(certificate = str_replace(certificate, "PG/A", "PGA"))

imdbtop1000v7 <- imdbtop1000v7%>%
  mutate(certificate = str_replace(certificate, "PGnrated", ""))

imdbtop1000v7 <- imdbtop1000v7%>%
  mutate(certificate = str_replace(certificate, 'NA', ""))

imdbtop1000v7 <- imdbtop1000v7 %>%
  mutate(across(where(is.character), ~ na_if(.,"")))

imdbtop1000v7 %>%
  count(genre)

imdbtop1000v7%>%
  slice_max(released_year, n = 1)

imdbtop1000v7%>%
  slice_min(released_year, n = 1)

imdbtop1000v8 <- imdbtop1000v7 %>%
  mutate(decade = case_when(released_year >= 1920 & released_year < 1930 ~ 1920,
                            released_year >= 1930 & released_year < 1940 ~ 1930,
                            released_year >= 1940 & released_year < 1950 ~ 1940,
                            released_year >= 1950 & released_year < 1960 ~ 1950,
                            released_year >= 1960 & released_year < 1970 ~ 1960,
                            released_year >= 1970 & released_year < 1980 ~ 1970,
                            released_year >= 1980 & released_year < 1990 ~ 1980,
                            released_year >= 1990 & released_year < 2000 ~ 1990,
                            released_year >= 2000 & released_year < 2010 ~ 2000,
                            released_year >= 2010 & released_year < 2020 ~ 2010,
                            released_year >= 2020 & released_year < 2030 ~ 2020))

head(imdbtop1000v8, n = 12)

imdbtop1000v8moviecountbd <- imdbtop1000v8%>%
  group_by(decade)%>%
  summarize(number_of_movies = n())

imdbtop1000v8moviecountbd

imdbtopcountbdviz <- ggplot(imdbtop1000v8moviecountbd,aes(decade, number_of_movies))+
  geom_bar(fill = "blue", stat = "identity")+
  theme_wsj()+
  labs(title = "Movie Count by Decade")

print(imdbtopcountbdviz)

imdb8.5highercountbd <- imdbtop1000v8%>%
  filter(imdb_rating >= 8.5)%>%
  group_by(decade)%>%
  summarize(number_of_movies = n())

ggplot(imdb8.5highercountbd, aes(decade, number_of_movies))+
  geom_bar(fill = "blue", stat= "identity")+
  scale_x_continuous(breaks = seq(1930, 2020, by =10))+
  theme_classic()+
  labs(title = "Movie Count by Decade with Imdb Rating of 8.5 or Higher")

imdbavgimdbbd <- imdbtop1000v8 %>%
  group_by(decade)%>%
  summarize(average_imdb_score = mean(imdb_rating))

idmbavgimdbbdviz <- ggplot(imdbavgimdbbd, aes(decade, average_imdb_score))+
  geom_bar(fill = "blue", stat = "identity")+
  scale_x_continuous(breaks = seq(1920, 2020, by = 10))+
  theme_classic()+
  labs(title = "Avg Imdb Rating by Decade", subtitle = "From top 1000 movies list with a minium rating of 7.6 and a maximum rating of 9.3")

print(idmbavgimdbbdviz)

imdbstar1count <- imdbtop1000v8 %>%
  group_by(star1)%>%
  summarize(number_of_movies = n())

imdbstar1count3 <- imdbstar1count%>%
  filter(number_of_movies >= 3)%>%
  select(star1, number_of_movies)%>%
  arrange(desc(number_of_movies))

print(imdbstar1count3)

imdbstar1count3top <- imdbstar1count3 %>%
  slice_max(number_of_movies, n = 6)

imdbstar1count3viz <- ggplot(imdbstar1count3top, aes(star1, number_of_movies))+
  geom_bar(fill = "blue", stat = "identity")+
  theme_classic()+
  labs(title = "Most Number of Movies Appeared in", subtitle = "from the imdb top 1000 movies dataset")

print(imdbstar1count3viz)

imdbstar2count <- imdbtop1000v8 %>%
  group_by(star2)%>%
  summarize(number_of_movies = n())

imdbstar2count2 <- imdbstar2count%>%
  filter(number_of_movies >= 2)%>%
  arrange(desc(number_of_movies))

print(imdbstar2count2)

imdbstar2count2top <- imdbstar2count%>%
  slice_max(number_of_movies, n = 3)

imdbstar2count2viz <- ggplot(imdbstar2count2top, aes(star2, number_of_movies))+
  geom_bar(fill = "blue", stat= "identity")+
  theme_classic()+
  labs(title = "Most Number of Movies Appearing in by a Supporting Actor", subtitle = "from the imdb top 1000 movies dataset")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

print (imdbstar2count2viz)

imdbdirectorcount <- imdbtop1000v8%>%
  group_by(director)%>%
  summarize(number_of_movies = n())

imdbdirectorcount3 <- imdbdirectorcount%>%
  filter(number_of_movies >= 3)%>%
  arrange(desc(number_of_movies))

print(imdbdirectorcount3)

imdbdirectorcount3top <- imdbdirectorcount3%>%
  slice_max(number_of_movies, n = 5)
  
imdbdirectorcount3viz <- ggplot(imdbdirectorcount3top, aes(director, number_of_movies))+
  geom_bar(fill = "blue", stat = "identity")+
  theme_classic()+
  labs(title = "Most Number of Movies Directed", subtitle = "from the imdb top 1000 movies dataset")

print(imdbdirectorcount3viz)

imdbdirectoravgimdb <- imdbtop1000v8%>%
  group_by(director)%>%
  summarize(average_imdb_rating = mean(imdb_rating))

imdbdirectoravgimdb1 <- imdbdirectoravgimdb %>%
  arrange(desc(average_imdb_rating))

print(imdbdirectoravgimdb1)
```
```


After running the count Frank Darabont leads the list with a 8.95 rating. The next thing I am going to figure out is which secondary supporting actor('star3' in the dataset), with a minimum of 2 movies, has the most appearances in the dataset. 
```{r finding the count of movies by star3 and filtering for 2 or more movies}

imdbstar3count <- imdbtop1000v8 %>%
  group_by(star3)%>%
  summarize(number_of_movies = n())

imdbstar3count2 <- imdbstar3count%>%
  filter(number_of_movies >= 2)%>%
  arrange(desc(number_of_movies))

print(imdbstar3count2)
```

After coming up with the count I see that Rupert Grint has the most appearances of any other secondary supporting actor with 5 appearances I can visualize the top 3 with ggplot...

```{r finding the top 3 counts and visualizing with ggplot}
imdbstar3counttop <- imdbstar3count2%>%
  slice_max(number_of_movies, n = 3)

imdbstar3countviz <- ggplot(imdbstar3counttop, aes(star3, number_of_movies))+
  geom_bar(fill = "blue", stat= "identity")+
  theme_classic()%>%
  labs(title = "Most Movies Appearing in by a Secondary Supporing Actor", subtitle = "from the imdb top 1000 movies dataset")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

print(imdbstar3countviz)
```

The next thing I would like to do is find the proportion of movies that appear on this list by genre, so first, I need to pull up a count of the genre...

```{r counting the genre column}
imdbtop1000v8%>%
  count(genre)
```

After pulling up the count I see that there are 202 genres, this is because the genres in the database are listed first with the main genre, and then further subcategorized with two other genres. In order to better determine the proportion of movies on this list by genre I will create a new column titled 'main_genre' and two other columns title 'alternate_genre1', 'alternate_genre2', I will use the separate function from the dplyr package to do this.

```{r separating the main_genre}

imdbtop1000v9 <- imdbtop1000v8%>%
  separate(genre, into = c("main_genre", "alternate_genre1", "alternate_genre2"), sep = ",")
```
After creating a new column I will now find the proportion of genres that appear in this dataset of top 1000 movies.

First I will determine the count of each 'main_genre'.



```{r finding the proportion of genres}
imdbtop1000v9genrecount <- imdbtop1000v9 %>%
  group_by(main_genre)%>%
  summarize(count = n())%>%
  mutate(percent = count/sum(count))

print(imdbtop1000v9genrecount)
```

Now I will visualize with ggplot...

```{r visualizing the proportion of genre from the dataset.}
genrecountproportionbar <- ggplot(imdbtop1000v9genrecount, aes(x = factor(main_genre), y = percent * 100, fill = main_genre))+
  geom_bar(stat = "identity")+
  labs(x = "main_genre", y = "percent", title = "Proportion of Genre", subtitle = "in the imdb top 1000 movies dataset")+
  scale_x_discrete(guide = guide_axis(n.dodge = 3))

print(genrecountproportionbar)  
```

After visualizing the genre count proportion it's clear that drama has the highest percentage of close to 30%. This could be visualized with a pie chart, however as shown below, with the amount of genre it's much easier to visualize with a bar chart.

```{r visualzing the proportion of genre from the dataset with a pie chart}
genrecountproportionpie <- ggplot(imdbtop1000v9genrecount, aes(x = "", y = percent * 100, fill = main_genre))+
  geom_col()+
  coord_polar(theta = "y")+
  labs(y = "Percent", title = "Proportion of Genre", subtitle = "in the imdb top 1000 movies dataset")

print(genrecountproportionpie)
```

After finding the proportion of genre in the top 1000 movies dataset, the next I am going to do is find the count of movies on this list in each year.

```{r finding the count of movies grouping by year}
moviescountbyyear <- imdbtop1000v9 %>%
  group_by(released_year)%>%
  summarize(count = n())

print(moviescountbyyear)


```

I will now sort it in descending order by count to see which years have the most appearances in this dataset.

```{r showing the top 15 counts}
moviescountbyyeartop <- moviescountbyyear%>%
  slice_max(count, n = 15)

print(moviescountbyyeartop)
```

After running the count 2014 has the most with 32, one of the things that sticks out is that of the top 15 in the movie count, only 1 comes from before the year 2000. 

```{r visualing the top 15 counts by release year}
moviescountbyyearviz <- ggplot(moviescountbyyeartop, aes(released_year, count))+
  geom_bar(stat = "identity", fill = "blue" )+
  theme_classic()+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

print(moviescountbyyearviz)


```

The next thing I am going to find out is what is the average runtime of movies by year in the dataset.

```{r attempting to find the average runtime by year}
movieruntimebyyear <- imdbtop1000v9 %>%
  group_by(released_year)%>%
  summarize(avg_runtime = mean(runtime))

print(movieruntimebyyear)
```

After trying to run this, I realize the runtime has 'min' also included which could be preventing me from getting the average runtime, so I will isolate the runtime without the 'min' portion so I will be able to aggregate the column.



```{r separting the min portion from each row into a new column}
imdbtop1000v10 <- imdbtop1000v9 %>%
  separate(runtime, into = c("runtime", "min"), sep = " ")
```

Now to change the name of the runtime column to indicate that it's in minutes...

```{r changing the column name to indicate that its shown in minutes}
colnames(imdbtop1000v10)[5] = "runtime_minutes"
 

``` 

Now I will remove the 'min' column since it's no longer needed...

```{r removing the min column from the dataset}
imdbtop1000v11 <- subset(imdbtop1000v10, select = - 6 )
```

Next I will try to aggregate the new runtime_minutes column to find the average runtime by year for movies in this dataset.

```{r attempting again to find the average runtime minutes by year}
movieruntimebyyear <- imdbtop1000v11 %>%
  group_by(released_year)%>%
  summarize(avg_runtime = mean(runtime_minutes))
```
After reading the error messages it's not able to aggregating the runtime column because it's not a numeric data type I will attempt to change the datatype. 

```{r changing runtime_minutes column from character data type to numeric data type}
imdbtop1000v11$runtime_minutes = as.numeric(as.character(imdbtop1000v11$runtime_minutes))

print(sapply(imdbtop1000v11, class))
```

After seeing that I successfully converted the column to numeric I will now retry to aggregate the runtime_minutes column to find the average runtime by year.

```{r attempting to find the average runtime by release year}
movieruntimebyyear <- imdbtop1000v11 %>%
  group_by(released_year)%>%
  summarize(avg_runtime = round(mean(runtime_minutes),2))
           
```
 I will now sort in descending to find the years with the maximum average runtime....
 
```{r sorting the avg runtime in descending order}
movieruntimebyyear %>%
  arrange(desc(avg_runtime))
```
 I can also visualize this with ggplot...
 
```{r visualizing the average runtime minutes by year}
movieruntimebyyearviz <- ggplot(movieruntimebyyear, aes(released_year, avg_runtime))+
  geom_point()+
  geom_line(aes(color = avg_runtime))+
  geom_smooth(method = "lm")+
  scale_x_discrete(breaks = seq(1920, 2030, by = 10))
  

print(movieruntimebyyearviz)
```
 
 In the visualizion is does appear that the over time the average runtime does seem to increase. 
 
    Well, that concludes my first data analysis project, over the course of this project I cleaned and validated the data including removing any unnecessary null values, making sure there weren't any duplicates, correcting any grammatical errors and any misspellings.
    
    After cleaning and validating the data, I then began to analyze the data by first finding out what decade has the most movies on this dataset by first creating a new 'decade' column and then sorting and visualizing the movie count. I found out that the decade containing the most movies in this dataset was the 2010's. I then got curious and wanted to find out if i were fo filter for movies with an imdb rating of 8.5 or higher would the movie count be different and indeed it was. The decade containing the most movies of a imdb rating of 8.5 or higher was 1990. 
      
    The next thing I wanted to figure out is what was the average imdb rating by decade. After aggregating the imdb_rating column by decade I was able to find out that the average imdb rating was highest (although not by much) in the years 2020 and 1920.
    
    Then, I wanted to find out which primary actor appeared most in the dataset, so I counted the primary actors and then found out that Tom Hanks appears the most with 12 movies. I then did the same with the supporting actor and found out Emma Watson had the most appearances with 7. AFter the supporting actor, I found out which director appeared the most and that was Alfred Hitchcock. 
    
    Then, I decided to figure out which director in the dataset had the highest imdb rating average and found that Frank Darabont has the imdb average rating with 8.95. After finding the highest average imdb rating by director, I found out that Rupert Grint has the most appearances on this list amongst secondary supporting actors.
    The next thing that I found out is that the largest proportion of genre in this dataset is 'Drama' with an almost 30% proportion. I then found out that 2014 has the most movies in this dataset with 32.
    
    Lastly, I wanted to find out the average runtime by year, in order to do this I first had to rename the 'runtime' column to 'runtime_minutes' and separate the 'min' into a new column and remove it entirely. I Then changed the datatype to numeric and found that from 1920 to about 1970 the average runtime trended upwards and overall the average runtime is higher than it was in the first 50 years in the dataset.
  


    
